---
title: "sacer63-genetic-diversity"
author: "Jenna McCullough"
date: "8/172023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(scales)
```

## Making a heterozygosity/pi dot chart 
This is following Devon DeRaad's scripts associated with DeRaad et al. 2023, Molecular Ecol. 
https://onlinelibrary.wiley.com/doi/10.1111/mec.17080


I have two datasets that I want to investigate, both using the sacer63 dataset.

1. **Subspecies**

2. **Individual**: all the samples from sacer63 are considered their own population for individual heterozygosity estimates 

```{r reading in output files}

#echo "GETTING GENETIC DIVERSITY OF SACER 63 subspecies"
#sacer63_subspecies.txt is a popmap that has all samples in their subspecies group  to get taxon level pi estimates.
#populations -P $src/populations_out/sacer63 -M $src/popmaps/sacer63_subspecies.txt --fstats -O $src/populations_out/pop_sacer63-subspeciesLumped -t 4

#echo "GETTING GENETIC DIVERSITY OF SACER 63 INDIVIDUALS"
# sacer67-individual.txt is a popmap that assigns samples to their own population to assess per individual heterozygosity estimates
#populations -P $src/populations_out/sacer63 -M $src/popmaps/sacer63-individual.txt -O $src/populations_out/pop_sacer63-individual -t 4

# then i had to separate the following tables into a "variant" and "all" file from the original. I also had to remove the top header and the # from the header line

#all subspecies 
pi.pops <-read.table("01_subspecies/populations.sumstats_summary-all.csv", header=T, sep=',')
popsmap<- read.table("sacer63-subspecies.txt", header=T, sep='\t')

# per sample 
pi.sample<-read.table("02_individual/populations.sumstats_summary-all.tsv", header=T, sep='\t')
samplemap <- read.table("sacer63-individuals.txt", header=T, sep='\t')

```


## all samples grouped by different popmaps
These following plots are the per sample run but plotted in a way that plots the individual
sample grouped by different metrics to explore patterns:

## grouped by subspecies
```{r pi sample1, echo=FALSE}
plotting.dftaxamap<-data.frame(sample=pi.sample$Pop.ID, het=pi.sample$Obs_Het, species=popsmap$pop)
#plot heterozygosity violin plots by viti 2 pop
ggplot(plotting.dftaxamap, aes(x=species, y=het))+ 
  #geom_violin(trim=FALSE)+
  geom_dotplot(binaxis='y', stackdir='center', dotsize = 1, alpha=.75)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


## Now I will add the population level estimates of Pi overlaid on top of the individual estimates of heterozygosity. 
Population levels of pi are on the right-hand side. The star is the pi value for the whole
population as defined by the populations stacks module. All the single points are
from when each sample was run as its own population. 

I will add colors to these in illustrator because i'm a lazy coder. 

```{r pi sample, echo=FALSE}

plotting.df<-data.frame(sample=pi.sample$Pop.ID, het=pi.sample$Obs_Het, species=popsmap$pop)
pi.pops$species<-pi.pops$Pop.ID
#plot heterozygosity violin plots
het.plot <-ggplot(plotting.df, aes(x=species, y=het))+ 
  #geom_violin(trim=FALSE)+
  geom_dotplot(binaxis='y', stackdir='center', dotsize = 1.9, alpha=0.75)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=12, face = "italic", color="black"),
        axis.text.y = element_text(angle = 45, hjust = 1, size = 12),
        legend.position = "none")+
  geom_point(pi.pops, mapping=aes(y=Pi), pch=8, cex=3)+
  labs(x="",y="heterozygosity")+
  scale_y_continuous(labels = label_comma(), sec.axis = sec_axis(trans = (~.*1), labels = label_comma(),name="Pi"))
plot(het.plot)
ggsave("het.pi.plot.pdf", het.plot, width = 6.5,height = 3,units = "in")

ggplot(plotting.df, aes(x=species, y=het))+ 
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  


```



